---
title: "preprocessing of the data from 3th and 4th run"
author: "L Perié"
date: "07/05/2019"
output: html_document
---

##1/ Convert the data file sent by Joost into one big file with all the samples. 

###1.1/ run 3
name coding is mouse_organ_month_celltype_rep
freq threshold=10
relative fraction threshold = 0
```{r, echo=FALSE}
#merge the data of one threshold together in one file
#threshold=0
#load package
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("plyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("reshape2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library(readr)
library(stringr)
#define working directory
knitr::opts_knit$set(root.dir = '~/Desktop/barcode mouse NL/analysis third sequencing rag+2/analyseleila' )
#define path
path="~/Desktop/barcode mouse NL/analysis third sequencing rag+2/analyseleila/half_axisbc_nonorm_0/"
#upload names of file in the folder
temp <- list.files(path="~/Desktop/barcode mouse NL/analysis third sequencing rag+2/analyseleila/half_axisbc_nonorm_0/")
#load data and generate merge dataframe
#create table
M <- data.frame()
#add first file
file_name <- str_sub(string = temp[1], start = 38, end = 48)
temp2 <- read.delim(paste0(path,temp[1]),header=TRUE)
colnames(temp2) <- c("tag",file_name)
assign( x = file_name, value = temp2, envir = .GlobalEnv)
M <- temp2
#add the rest of the files
for (i in 2:length(temp)) {
  file_name <- str_sub(string = temp[i], start = 38, end = 48)
  temp2 <- read.delim(paste0(path,temp[i]),header=TRUE)
  colnames(temp2) <- c("tag",file_name)
  M <- merge(M, temp2,all=TRUE)
  assign( x = file_name, value = temp2, envir = .GlobalEnv)
  }
M[is.na(M)]<-0 # assign zero to Nas
#substract difference in names
colnames(M) <- gsub("a_","a" , colnames(M))
colnames(M) <- gsub("b_","b" , colnames(M))
write.table(M, "bcm run 3 freq10frac0_allsamples.txt",row.names=TRUE,col.names=NA, sep="\t")
```


###1.2/ run 4

```{r, echo=FALSE}
#merge the data from sequencing run 4th of different samples together in one file
#load package
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("plyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("reshape2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library(readr)
library(stringr)
#define working directory
knitr::opts_knit$set(root.dir = '/Users/leila/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2' )
#define path
path="/Users/leila/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/cleanedfinal_halfsams_freq10relfrac0_/"
#upload names of file in the folder
temp <- list.files(path="/Users/leila/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/cleanedfinal_halfsams_freq10relfrac0_")
#load data and generate merge dataframe
#create table
M <- data.frame()
#add first file
file_name <- str_sub(string = temp[1], start = 38, end = 50)
temp2 <- read.delim(paste0(path,temp[1]),header=TRUE)
colnames(temp2) <- c("tag",file_name)
assign( x = file_name, value = temp2, envir = .GlobalEnv)
M <- temp2
#add the rest of the files
for (i in 2:length(temp)) {
  file_name <- str_sub(string = temp[i], start = 38, end = 50)
  temp2 <- read.delim(paste0(path,temp[i]),header=TRUE)
  colnames(temp2) <- c("tag",file_name)
  M <- merge(M, temp2,all=TRUE)
  assign( x = file_name, value = temp2, envir = .GlobalEnv)
}
M[is.na(M)]<-0 # assign zero to Nas
#substract difference in names
colnames(M) <- gsub("a_k","a" , colnames(M))
colnames(M) <- gsub("b_k","b" , colnames(M))
colnames(M) <- gsub("a_","a" , colnames(M))
colnames(M) <- gsub("b_","b" , colnames(M))
colnames(M) <- gsub("ae","a" , colnames(M))
colnames(M) <- gsub("be","b" , colnames(M))
write.table(M, "bcm run 4 freq10frac0_allsamples.txt",row.names=TRUE,col.names=NA, sep="\t")
```


##2/ merge file run 3 and run 4

```{r, echo=FALSE}
d3 <- read.delim("~/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/bcm run 3 freq10frac0_allsamples-repeatedsamples.txt")
d4 <- read.delim("~/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/bcm run 4 freq10frac0_allsamples.txt")
d3 <- d3[,-1]
d4 <- d4[,-1]
D <-merge(d3,d4,by="tag",all=TRUE) 
D[is.na(D)]<-0 # assign zero to Nas
write.table(D, "bcm run 3 and 4 freq10frac0_allsamples.txt",row.names=TRUE,col.names=NA, sep="\t")

```


##3/renorm
```{r, echo=FALSE}
#norm
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("plyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("reshape2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library(readr)
library(stringr)
#load data of all samples from run 3 and 4
M <- read.delim("~/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/bcm run 3 and 4 freq10frac0_allsamples.txt")
M <- M[,-1]

#normalization to 1
rownames(M) <- M[,1]
data <- M[,-1]
norm.data <- apply(data,2, function(x) (x/sum(x)))
write.table(norm.data, "BCM run 3 and 4 freq10frac0_allsamples norm 1.txt",row.names=TRUE,col.names=NA, sep="\t")

```


##4/ apply read threshold of 0.003
```{r, echo=FALSE}
#fraction threshold is 0.003 based on MEF data
T3=0.003
d3 <- read.delim("~/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/BCM run 3 and 4 freq10frac0_allsamples norm 1.txt")
M3 <- d3[,-1]
M3[M3<T3] <-0 
M3 <- cbind(d3[,1],M3)
names(M3)[1]<- "tag"
write.table(M3,"BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003.txt",row.names=TRUE,col.names=NA, sep="\t")
```


##5/renorm
```{r, echo=FALSE}
#renorm
#load data 
M <- read.delim("~/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003.txt")
M <- M[,-1]

#renormalization to 1
rownames(M) <- M[,1]
data <- M[,-1]
norm.data <- apply(data,2, function(x) (x/sum(x)))
write.table(norm.data, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm.txt",row.names=TRUE,col.names=NA, sep="\t")

```


##5/change format to long

```{r, echo=FALSE}
#change format of the data and generate a table with duplicates
library("reshape2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
d1 <- read.delim("BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm.txt")
names(d1)[1]<- "tag"
d2 <- melt(d1, id.vars="tag")
d2<- d2[which(d2$value>0),]
d3 <- cbind( d2, ldply(strsplit(as.character(d2$variable), "_"), identity) ) # split the identifier column
names(d3) <- c("tag","id","var","mouse","month","organ","type","rep") #give name to categories
d3<- d3[which(d3$var>0),]
da <- d3[which(d3$rep == 'a'),]#data of rep a
da <- da[,!colnames(da)=="rep"]#delete colonn rep
da <- da[,!colnames(da)=="id"]
row.names(da) <- NULL #delete the name of the row
dimnames(da) [[2]][2]<- "vara"#rename the colonn var en vara
db <- d3[which(d3$rep == 'b'),]#idem with b
row.names(db) <- NULL
db <- db[,!colnames(db)=="rep"]
db <- db[,!colnames(db)=="id"]
dimnames(db) [[2]][2]<- "varb"
dab <- merge(da,db,all=T)#create one table with a and b on the same row
dab[is.na(dab)]<-0 #delete na
write.table(dab, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm a vs b.txt",row.names=TRUE,col.names=NA, sep="\t")
```

##6/ sum duplicates
```{r, echo=FALSE}
#make a file with sum of duplicates
#load data with duplicates
d <- read.delim("BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm.txt")
names(d)[1]<- "tag"
#add duplicates
D <- as.data.frame(matrix(0, ncol = (dim(d)[2]-1)/2, nrow = dim(d)[1]))
j=0
for (i in seq(2,dim(d)[2],2)) {
  j=j+1
  D[,j] <- (d[,i]+d[,i+1])
}
c <- colnames(d[,-1])  
c <- gsub("_a","" , c)
c <- gsub("_b","" ,c)
names(D) <- unique(c)
D <- cbind(d[,1],D)
names(D)[1] <- "tag"
write.table(D, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm repsum.txt",row.names=TRUE,col.names=NA, sep="\t")

```

##7/ renorm to cell numbers 
 
```{r, echo=FALSE}
#load data
#I have deleted from the cell number given by Joost the samples that were deleted from run 3 as they are repeated in run 4 (1_BM_15,2_BM_15 and 2_4_BL_L, 2_5_BL_M and 1_5_BL_M)
numb <- read.delim("cellnumbrun3and4.txt")
d <- read.delim("BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm.txt")
names(d)[1]<- "tag"
#renorm D to cell number
rownames(d) <- d[,1]
data <- d[,-c(1)]
norm.data <- data.frame(matrix(NA, ncol=dim(data)[2], nrow=dim(data)[1]))
for (i in seq(1:dim(data)[2])) {
  name <- colnames(data)[i]
  name <- substring(name,2)
  n <- numb[which(numb[,1]==name),2]
  norm.data[,i] <-data[,i]/sum(data[,i])*n/2
}
rownames(norm.data) <- rownames(d)
colnames (norm.data) <-colnames (data) 
write.table(norm.data, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb.txt",row.names=TRUE,col.names=NA, sep="\t")
```

##8/sum rep cell numbers
```{r, echo=FALSE}
d <- read.delim("BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb.txt")
N <- d[,-1]
D <- as.data.frame(matrix(0, ncol = dim(N)[2]/2, nrow = dim(N)[1]))
j=0
for (i in seq(1,dim(N)[2],2)) {
    j=j+1
    D[,j] <- (N[,i]+N[,i+1])
}
c <- colnames(N)  
c <- gsub("_a","" , c)
c <- gsub("_b","" ,c)
names(D) <- unique(c)
rownames(D) <- d[,1]
write.table(D, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb repsum.txt",row.names=TRUE,col.names=NA, sep="\t")
```

##9/ convert to long
```{r, echo=FALSE}
#change format of the data and generate a table with duplicates
library("reshape2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
d1 <- read.delim("BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb.txt")
names(d1)[1] <- "tag"
d2 <- melt(d1, id.vars="tag")
d2<- d2[which(d2$value>0),]
d3 <- cbind( d2, ldply(strsplit(as.character(d2$variable), "_"), identity) ) # split the identifier column
names(d3) <- c("tag","id","var","mouse","month","organ","type","rep") #give name to categories
d3<- d3[which(d3$var>0),]
da <- d3[which(d3$rep == 'a'),]#data of rep a
da <- da[,!colnames(da)=="rep"]#delete colonn rep
da <- da[,!colnames(da)=="id"]
row.names(da) <- NULL #delete the name of the row
dimnames(da) [[2]][2]<- "vara"#rename the colonn var en vara
db <- d3[which(d3$rep == 'b'),]#idem with b
row.names(db) <- NULL
db <- db[,!colnames(db)=="rep"]
db <- db[,!colnames(db)=="id"]
dimnames(db) [[2]][2]<- "varb"
dab <- merge(da,db,all=T)#create one table with a and b on the same row
dab[is.na(dab)]<-0 #delete na
write.table(dab, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb a vs b.txt",row.names=TRUE,col.names=NA, sep="\t")
```

##10/ duplicate filtering (ab filtering)

```{r, echo=FALSE}
dab <- read.delim("BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb a vs b.txt")
dab <- dab[,-1]
z1 <- dab[which(dab$organ!="BL"),]
z1 <- z1[which(z1$vara>0 & z1$varb>0),]
write.table(z1, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb a vs b not blood samples-abfilt.txt",quote=F, sep="\t", row.names=F )

#plot new self/self after abfiltering
qplot(asinh(vara), asinh(varb), data=z1) + facet_wrap(~paste( mouse,month, type, organ, sep = " _ "))

```

##11/ convert to wild
```{r, echo=FALSE}
#transform format to wild format of z1
z1[,8] <-paste(z1$mouse,z1$month,z1$organ,z1$type, sep="_")
x1 <- z1[,c(1,6,7,8)]
xa <- x1[,-3]
names(xa) <- c("tag","var", "id")
xa$id <-paste(xa$id,"a", sep="_")
xb <- x1[,-2]
names(xb) <- c("tag","var", "id")
xb$id <-paste(xb$id,"b", sep="_")
x1<- rbind(xa,xb)

x1 <- reshape(x1,direction="wide", timevar="id", idvar="tag" )
x1[is.na(x1)] <- 0
names(x1) <- sub("var.","", colnames(x1))
write.table( x1, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb not blood samples-abfilt wilde.txt",quote=F, sep="\t", row.names=F )
```

## 12/ repsum and convert again to wild
```{r, echo=FALSE}
z1 <- read.delim("BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb a vs b not blood samples-abfilt.txt")
z1 <- ddply(z1, c("tag","mouse","month", "type","organ", "vara","varb"), summarize, sum=vara+varb)#sum replicates
z1[,9] <-paste(z1$mouse,z1$month,z1$organ,z1$type, sep="_")
x1 <- z1[,c(1,8, 9)]
names(x1) <- c("tag","var", "id")
x1 <- reshape(x1,direction="wide", timevar="id", idvar="tag" )
x1[is.na(x1)] <- 0
names(x1) <- sub("var.","", colnames(x1))
write.table(x1, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb repsum not blood samples-abfilt.txt",row.names=TRUE,col.names=NA, sep="\t")
```

##13/file sequences of the barcodes
```{r, echo=FALSE}
#merge the split bc from run 4
#threshold=0
#load package
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("plyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("reshape2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library(readr)
library(stringr)
#define working directory
knitr::opts_knit$set(root.dir = '/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2' )
#define path
path="/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/splitbcs_/"
#upload names of file in the folder
temp <- list.files(path="/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/splitbcs_")
#load data and generate merge dataframe
#create table
temp2 <- read.table(paste0(path,temp[1]),header=TRUE, sep = "\t", fill = TRUE, colClasses=c("insDJ"="character"))
name <- names(temp2)
M <- temp2
M$tag <- rownames(M)
rownames(M) <- NULL
#length(temp)
#add the rest of the files
for (i in 2:length(temp)) {
  temp2 <- read.table(paste0(path,temp[i]),header=TRUE, sep = "\t", fill = TRUE, colClasses=c("insDJ"="character"))
  temp2$tag <- rownames(temp2)
  
  rownames(temp2) <- NULL
  M<- merge(M, temp2,all=TRUE)
}
write.table(M, "splitbc_run4_allsamples.txt",row.names=FALSE, sep="\t")
#issue with some NAs that duplicate columns check this with length(unique(M)) that should be egal to dim(M)[1]. To know how many duplicates du M[duplicated(M$tag),] --> I have deleted the 5 duplicates by end. 

#merge with run 3
run3 <- read.delim("splitbcs_run3_freq10frac0.txt")
run4<- read.delim("splitbc_run4_allsamples_woNA.txt")
name <- names(run4)
M <- merge(run3,run4,by=name,all=TRUE)
write.table(M, "splitbc_run3and4_allsamples.txt",row.names=FALSE, sep="\t")

```

##14/ merge probabilty file from yuval with the data file
```{r,echo=FALSE, warning=FALSE,message=FALSE, cache=FALSE}
#preprocess Yuval's file
setwd("/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/yuval")
Pgen <- read.delim("/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/yuval/run34_clean_pgen_likelihood.txt",sep="\t")
Pgen <- Pgen[,-c(1,3)]
names(Pgen)[1] <- "tag"

hist(log10(Pgen[,2]),xlab="sequence probability(log10)")

Dall <- read.delim("/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb repsum not blood samples-abfilt.txt",sep="\t")
Dall <- Dall[,-1]
names(Dall)[1] <- "tag"

setwd("/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila")

D <- merge(Dall,Pgen, by="tag")
D[is.na(D)]<-0

write.table(D, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb repsum not blood samples-abfilt probability yuval.txt",row.names=FALSE,col.names=TRUE, sep="\t")

```

#15/ super mouse more bc on the axis removed
```{r, echo=FALSE}
d <- read.delim("/Users/lperie/Documents/Leila/NL lab/projet NL/barcode mouse NL/analysis third and fourth sequencing rag+2/analyseleila/BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb repsum not blood samples-abfilt probability yuval.txt",sep="\t")

LSK <- d[, grep("_LSK", colnames(d))]

x1 <- d[, grep("X1_", colnames(d))]
x1 <- x1[,-grep("_L", colnames(x1))]
x1 <- x1[, -grep("_DC", colnames(x1))]
X1_15_BM_LSK<- LSK[, grep("X1_", colnames(LSK))]
x1 <- cbind(x1,X1_15_BM_LSK)
x1 <- cbind(x1, d$Pgen)
x1 <-x1[rowSums(x1[,-5])>0,]
x1 <- x1[which(x1$`d$Pgen`!=0),]

M <- cbind(x1,1)
names(M) <- sub("X1_","", colnames(x1))
names(M)[6] <- "mouse"

x1 <- d[, grep("X2_", colnames(d))]
x1 <- x1[,-grep("_L", colnames(x1))]
x1 <- x1[, -grep("_DC", colnames(x1))]
X2_15_BM_LSK<- LSK[, grep("X2_", colnames(LSK))]
x1 <- cbind(x1,X2_15_BM_LSK)
x1 <- cbind(x1, d$Pgen,2)
x1 <-x1[rowSums(x1[,-5])>0,]
x1 <- x1[which(x1$`d$Pgen`!=0),]
names(x1) <- sub("X2_","", colnames(x1))
names(x1)[6] <- "mouse"
  
M <- rbind(M,x1)

x1 <- d[, grep("X3_", colnames(d))]
x1 <- x1[,-grep("_L", colnames(x1))]
x1 <- x1[, -grep("_DC", colnames(x1))]
X3_15_BM_LSK<- LSK[, grep("X3_", colnames(LSK))]
x1 <- cbind(x1,X3_15_BM_LSK)
x1 <- cbind(x1, d$Pgen,3)
x1 <-x1[rowSums(x1[,-5])>0,]
x1 <- x1[which(x1$`d$Pgen`!=0),]
names(x1) <- sub("X3_","", colnames(x1))
names(x1)[6] <- "mouse"

M <- rbind(M,x1)

x1 <- d[, grep("X4_", colnames(d))]
x1 <- x1[,-grep("_L", colnames(x1))]
x1 <- x1[, -grep("_DC", colnames(x1))]
X4_15_BM_LSK<- LSK[, grep("X4_", colnames(LSK))]
x1 <- cbind(x1,X4_15_BM_LSK)
x1 <- cbind(x1, d$Pgen,4)
x1 <-x1[rowSums(x1[,-5])>0,]
x1 <- x1[which(x1$`d$Pgen`!=0),]
names(x1) <- sub("X4_","", colnames(x1))
names(x1)[6] <- "mouse"

M <- rbind(M,x1)
write.table(M, "BCM run 3 and 4 freq10frac0_allsamples norm 1_0.003 renorm norm cell numb repsum not blood samples-abfilt probability yuval super mouse.txt",row.names=FALSE, sep="\t")

```



